Missing links and the topological robustness of food webs

### 2023-02-10 manuscript/ms_C3_v2.tex
updated ms

\usepackage{lineno}
\usepackage {amsmath}
\setlength\parindent{24pt}
\usepackage{setspace}\doublespacing
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\begin{document}
\maketitle

\begin{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  Undersampling can lead to missing trophic interactions in recorded
  food webs, with potential consequences for the perceived functioning
  and stability of the food webs. Undersampling can be compensated for
  by using food web models such as the allometric diet breadth model
  (ADBM) to predict missing links. Simultaneously, models might predict
  links which cannot occur, i.e., false positives.
\item
  Previous research shows that (i) food web robustness (the inverse of
  the number of secondary extinctions occuring due to primary
  extinctions) increases with connectance (the number of realised
  trophic links divided by the number of possible links), and (ii) that
  predicted food webs usually have greater connectance than observed
  ones. Thus we expect that predicted food webs are more robust than
  observed ones. This expectation has never, to our knowledge, been
  tested, nor has the effect size been quantified.
\item
  We fill this research gap by comparing the robustness of observed food
  webs to the robustness of food webs predicted by a model (the ADBM)
  that can account for missing links, though can also make false
  positives. We did this for 12 different food webs from a wide variety
  of ecosystems.
\item
  We found, as expected, that the predicted food webs were more robust
  than the observed food webs, and this can be attributed to the higher
  connectance of the predicted food webs. On average, for every one unit
  of increase in connectance, we found the food webs to be robust by
  0.52 units and 0.04 units for the most connected and the random
  species extinction scenarios respectively while no effect in the least
  connected species extinction scenario.
\item
  These results show that undersampling can lead to large underestimates
  of food web robustness that can be compensated for by filling in
  missing links with food web models. Nevertheless, increased
  connectance may contribute to lower dynamical stability, and so it
  would be interesting to compare the dynamical stability of observed
  and predicted food webs, as well as the topological stability that we
  have focused on.
\end{enumerate}
\end{abstract}

@@

Anthropogenic changes such as climate change and habitat destruction are
a threat to biodiversity, and can lead to food web collapse (Ullah et
al. 2018). This food web collapse is due to the cascades of secondary
extinctions in a food web because of the primary loss of species, for
example due to habitat destruction and climate change (Pimm et al. 2006;
J. A. Thomas et al. 2004; C. D. Thomas et al. 2004). Therefore, research
focused on cascading secondary extinctions also known as `community
viability analysis' have been performed extensively in the past few
decades to quantify how robust are food webs to species extinction
(Jennifer A. Dunne, Williams, and Martinez 2002b; Jennifer A. Dunne and
Williams 2009; Berg et al. 2011; Ebenman, Law, and Borrvall 2004;
Ebenman and Jonsson 2005). And it has been shown that the rate of
collapse of a food web is dependent on its structure and complexity
(Jennifer A. Dunne, Williams, and Martinez 2002b; Jennifer A. Dunne and
Williams 2009).

Simulation of primary species loss has been conducted in observed food
webs and model food webs from terrestrial and aquatic ecosystems where
robustness was measured in terms of secondary extinctions (Jennifer A.
Dunne, Williams, and Martinez 2002b; Jennifer A. Dunne and Williams
2009). These studies have shown that the robustness of the food webs
increases with food web connectance. Also, the removal of the most
connected species cause considerably more secondary extinctions than the
random removals of species (Jennifer A. Dunne, Williams, and Martinez
2002a; Sol'e and Montoya 2001). Simulation studies like these which
investigate the impact of primary extinctions in a food web to quantify
robustness based on its topological structure provide an alternate
solution to canonical experiments in natural ecosystems which are not
possible or very difficult to conduct (Jennifer A. Dunne and Williams
2009).

Along with quantifying food web robustness based on its topological
structure, studies such as Williams (2008), Brose, Williams, and
Martinez (2006) and Martinez, Williams, and Dunne (2006) have quantified
robustness based on the abundance dynamics of a food web. The
topological approach of quantifying a food web robustness only requires
the food web structure whereas the dynamical approach not only requires
the food web structure but also the temporal dynamics of abundance of
species in that food web. For example: Williams (2008) combined models
of network structure with models of bioenergetic dynamics to study the
role of food web topology and nonlinear dynamics on species coexistence
in complex ecological networks.

A key assumption of the observed food webs is that they are very well
sampled i.e.~all the links that in reality can occur are represented.
However, it is known that not all food webs are very well sampled and
then do not represent all of the feeding links that occur (Caron et al.
2022; Patonai and Jord'an 2017; Jordano 2016). Some rare trophic links
require more sampling effort as compared to others whereas some trophic
links remain unobserved because of linkage constraints irrespective of
sufficient sampling effort (Jordano 2016). Previous studies such as
Caron et al. (2022) and Gupta, Furrer, and Petchey (2022) have shown
that the predicted food webs from these models usually have greater
connectance than the observed ones. Therefore, one solution to
compensate for undersampling is to use a food web model such as the
Allometric Diet Breadth Model (ADBM) to predict the missing links, and
to then measure the robustness of the predicted food web. ADBM is a
mechanistic model constructed using foraging rules based on the body
sizes of prey and predator where trophic interactions satisfying those
rules would be predicted by the model which are perhaps not observed
because those interactions are rare. However, this solution is not
infallible, as it is likely that the food web model might still miss
some links, and also may predict some links that could not, in fact
occur.

In our study, we investigate the topological robustness of the ADBM
predicted food webs and compare it to that of the observed food webs,
and quantify the effect of undersampling i.e overestimation of
connectance on the robustness of these predicted food webs. We expect
that the ADBM predicted food webs to be more robust as compared to the
observed food webs, and for the greater robustness to be related to the
amount by which the ADBM overestimates connectance. We do this by
simulating primary species loss in 12 food webs predicted from the ADBM
to quantify the secondary loss of extinctions. We use three different
approaches of species removal: (i) most connected species, (ii) random
species and (iii) least connected species to understand if the outcome
varies among these approaches.

@@

In the upcoming sections, we present a detailed account of the
implementation of simulation of primary extinctions for three different
extinction scenarios on 12 food webs predicted by the ADBM from a wide
variety of ecosystems and compute the resultant secondary extinctions.
We then compute a robustness metric to quantify the robustness of those
predicted food webs and compare them against that of the observed food
webs.

@@

The allometric diet breadth model (ADBM) is based on optimal foraging
theory, specifically the contingency model (MacArthur and Pianka 1966).
We chose this model because the ADBM can likely predict missing links in
the predicted food webs because it consistently overestimated
connectance in the predicted food webs as shown by our study in Gupta,
Furrer, and Petchey (2022). The ADBM predicts the set of prey species a
consumer should feed upon to maximise its rate of energy intake (Petchey
et al. 2008). The foraging variables used in the model are the energy
content of prey, handling times of the predator on prey, space clearance
rate i.e.~how fast a predator searches space, and prey densities. All of
these variables are derived from the allometric scaling relationship
using the body sizes of species. More details on the foraging rules
defined in the ADBM and ADBM's predictive power across different food
webs can be found in Petchey et al. (2008).

@@

The observed food webs that we fit the ADBM to belong to marine,
freshwater and terrestrial ecosystems (Table \ref{fig:tab_1}). These
food webs contain primary producers, herbivores, carnivores, parasites,
and parasitoids and also contain various types of feeding interactions,
including predation, herbivory, bacterivory, parasitism and pathogenic.
The observed connectance of these food webs varies from 0.03 to 0.24 and
the number of species varies from 29 to 239 species.

The goodness of fit of the ADBM's predictions depends on the interaction
types in the food webs with predictions that are more size-structured
such as aquatic and herbivory interactions being better predicted when
compared to less size-structured ones such as parasitoid and terrestrial
herbivory ones (Petchey et al. 2008).

@@

\caption{\label{tab:unnamed-chunk-1}\label{fig:tab_1}Information about the food webs predicted using the ADBM.}
\centering
\resizebox{\linewidth}{!}{
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}|>{\raggedright\arraybackslash}p{8em}|l|l|l|l|>{}p{8em}|>{}p{8em}|>{}p{8em}}
\hline
Common food web name (Original Publication) & Predation matrix source & General ecosystem & Number of species & Observed connectance & 95\% prediction interval of predicted connectance  (Gupta et al. 2022)\\
\hline
Benguela Pelagic (Yodzis 1998) & Brose et al. (2005) & Marine & 30 & 0.21 & 0.26 - 0.59\\
\hline
Broadstone Stream (taxonomic aggregation) (Woodward and Hildrew 2001; Woodward
et al. 2005) & Brose et al. (2005) & Freshwater & 29 & 0.19 & 0.18 - 0.72\\
\hline
Broom (Memmott et al. 2000) & Brose et al. (2005) & Terrestrial & 60 & 0.03 & 0.12 - 0.89\\
\hline
Capinteria (Lafferty et al. 2006) & Hechinger et al. (2011) & Marine (Salt Marsh) & 88 & 0.08 & 0.11 - 0.80\\
\hline
Caricaie Lakes (Cattin et al. 2004) & Brose et al. (2005) & Freshwater & 158 & 0.05 & 0.11 - 0.81\\
\hline
Grasslands (Dawah et al. 1995) & Brose et al. (2005) & Terrestrial & 65 & 0.03 & 0.03 - 0.44\\
\hline
Mill Stream (Ledger, Edwards, Woodward unpublished) & Brose et al. (2005) & Freshwater & 80 & 0.06 & 0.08 - 0.60\\
\hline
Skipwith Pond (Warren 1989) & Brose et al. (2005) & Freshwater & 71 & 0.07 & 0.17 - 0.90\\
\hline
Small Reef (Opitz 1996 Table 8.6.2) & Alyssa R. Cirtwill and Anna Eklöf (2018) & Marine (Reef) & 239 & 0.06 & 0.07 - 0.66\\
\hline
Tuesday Lake (Jonsson et al. 2005) & Brose et al. (2005) & Freshwater & 73 & 0.08 & 0.09 - 0.57\\
\hline
Ythan (Emmerson and Raffaelli 2004) & Alyssa R. Cirtwill and Anna Eklöf (2018) & Marine (Estuarine) & 85 & 0.04 & 0.13 - 0.84\\
\hline
Broadstone Stream (size aggregation) (Woodward
et al. 2010) & Guy Woodward. (2021) & Freshwater & 29 & 0.24 & 0.25 - 0.47\\
\hline
\end{tabular}}
\end{table}
\end{landscape}
\restoregeometry

@@

The ADBM was parameterised using approximate Bayesian computation (ABC)
where a set of parameter values were sampled from the prior
distributions. That set of parameter values was either accepted or
rejected based on how close the predicted food web is to the observed
food web using an accuracy metric -- true skill statistic (TSS). The
accepted parameter values then formed a posterior distribution. Further,
prediction intervals of the true skill statistic and connectance of the
predicted food webs were computed. In our study, we considered food webs
where the predicted connectance lay within the 95\% prediction interval.
A detailed explanation of the parameterisation method can be found in
Gupta, Furrer, and Petchey (2022).

@@

We implemented the primary species removal method from Jennifer A. Dunne
and Williams (2009) by sequentially removing species using one of the
three criteria: removal of (i) the most-connected species, (ii) the
least-connected species and (iii) randomly chosen species. The
most-connected and least-connected criteria are based on the degree
(i.e.~the total number of links to resources and from consumers) of
species. We considered these mentioned criteria because the random
extinction scenario takes into account all the theoretically possible
extinction sequences of species that can occur in a food web and the
extinction of most-connected species and least-connected species takes
into account the two opposite extreme scenarios. These extinction
scenarios have been widely used in studying species extinctions and
collapse of food webs and other networks (Jennifer A. Dunne, Williams,
and Martinez 2002b; Sol'e and Montoya 2001; J. Dunne, Williams, and
Martinez 2004; Jennifer A. Dunne and Williams 2009; Albert and Barab'asi
2002).

Given a primary removal of species in a food web if any remaining
species lost all of their resource species, or any cannibalistic species
lost all of their resource species except the cannibalistic links, they
are removed from the web and a secondary extinction was recorded.
Secondary extinctions may cause further secondary extinctions, which
were also checked for and recorded. Once no more secondary extinctions
occurred, then another primary extinction was made, of the next
appropriate species depending on the extinction scenario. This process
was carried out until all the species were extinct from the food web.

The robustness (R) of a food web was defined as the proportion of
species subjected to primary removals that resulted in a set of
extinction (i.e.~primary removals plus secondary extinctions) of some
specified proportion of the species. In our study, we use \(R_{50}\),
the number of primary extinctions divided by the total number of
species, which results in at least 50\% of total species loss (Jennifer
A. Dunne, Williams, and Martinez 2002b; J. Dunne, Williams, and Martinez
2004; Jonsson et al. 2015; Jennifer A. Dunne and Williams 2009).
Therefore, if primary extinctions never cause any secondary extinctions,
the food web is maximally robust and (\(R_{50} = 0.50\)). Whereas in a
minimally robust community (\(R_{50} = 1/S\)), the first primary
extinction causes a cascade of secondary extinctions of at least nearly
half of the species in the food web (i.e.~at least \(S/2 - 1\)).

@@

First, we simulated primary species loss in food webs predicted by the
ADBM which had the maximum true skill statistics and compared it to
primary species loss in observed food webs. Second, to take into account
the uncertainty in robustness in the ADBM predicted food webs we
simulated primary species loss and thereby computed robustness for all
the ADBM predicted food webs corresponding to the 95\% prediction
interval of the predicted connectance. In the case of the random
extinction scenario, we simulated 1000 random extinction sequences in a
single ADBM predicted food web.

\hypertarget{statistical-analysis}{%
\subsection{Statistical analysis}\label{statistical-analysis}}

In the random extinction scenario, we computed robustness \(R_{50}\) for
all 1000 independent random extinction sequences and calculated the
median as a summary statistics to quantify the average robustness of a
single food web. To quantify the effect of undersampling
i.e.~overestimation of connectance we compute the ratio of the
difference in normalised robustness between the ADBM predicted food webs
and observed food webs to the difference in their normalised
connectance, where normalisation was performed by dividing the variables
by their maximum possible values (I.e 0.5 for \(R_{50}\) and 1 for
connectance). However, we did not perform any statistical significance
test because we work with simulated food webs and therefore the p-values
of these tests are influenced by the number of model simulations (White
et al. 2014).

@@

We first present the secondary extinction curves of the ADBM predicted
food webs which had the maximum true skill statistics and the observed
food webs for 12 food webs under three different extinction scenarios.
We then compare the robustness of all the ADBM predicted food webs
within the 95\% prediction interval against that of the observed food
webs to take into account uncertainty in the robustness across food webs
predictions. Finally we quantify the effect of overestimation of
connectance on the difference in their robustness estimates.

@@

In the most-connected extinction scenario, the cumulative secondary
extinction curves started to rise steeply at a lower number of primary
species removal in the observed food webs as compared to the ADBM
predicted food webs for nine food webs (Fig. \ref{fig:fig_r1} (a, c, d,
e, f, g, h, j, l)). However there were higher number of cumulative
secondary extinctions occurring in the ADBM predicted food webs when
compared to that of the observed food webs at a high number of primary
species removal (Fig. \ref{fig:fig_r1} (a, f, g, h, k)). In the Skipwith
Pond food web, there were no secondary extinctions for any number of
primary removal of species (Fig. \ref{fig:fig_r1} (i)), whereas in the
Broadstone Stream (taxonomic aggregation) food web the same was true
only for the observed food web but in the ADBM predicted food web there
was a steep rise in the cumulative secondary extinctions (Fig.
\ref{fig:fig_r1} (b)).

@@

\caption{\label{fig:fig_r1} Cumulative secondary extinctions of species resulting from the primary \textbf{removals of the most connected species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species.}\label{fig:unnamed-chunk-2}
\end{figure}

In Fig. \ref{fig:fig_r2}, we present the cumulative secondary
extinctions in the ADBM predicted and the observed food webs for five
(out of 1000) independent random extinction sequences to show
qualitative differences between their secondary extinction curves. The
secondary extinctions in the ADBM predicted food webs were more abrupt
than that in the observed food webs.

@@

\caption{\label{fig:fig_r2} Cumulative secondary extinctions of species resulting from the primary \textbf{removals of random species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs for five independent random extinction sequences. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species.}\label{fig:unnamed-chunk-3}
\end{figure}

>>>>

\usepackage{lineno}
\linenumbers
\usepackage {amsmath}
\setlength\parindent{24pt}
\usepackage{setspace}\doublespacing
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\begin{document}
\maketitle

\begin{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  Undersampling can lead to missing trophic interactions in recorded
  food webs, with potential consequences for the perceived functioning
  and stability of the food webs. Undersampling can be compensated for
  by using food web models such as the allometric diet breadth model
  (ADBM) to predict missing links. Simultaneously, models might predict
  links which cannot occur, i.e., false positives.
\item
  Previous research shows that (i) food web robustness (the inverse of
  the number of secondary extinctions occurring due to primary
  extinctions) increases with connectance (the number of realised
  trophic links divided by the number of possible links), and (ii) that
  model predicted food webs usually have greater connectance than
  observed ones. Thus, we expect that predicted food webs are more
  robust than observed ones. However, this expectation has never, to our
  knowledge, been tested, nor has the effect size, i.e.~the difference
  in robustness of predicted and observed food web with respect to the
  difference in their connectance, been quantified.
\item
  We fill this research gap by comparing the robustness of observed food
  webs to the robustness of food webs predicted by a model (the ADBM)
  that can account for missing links but does not predict all links
  correctly. We did this for 12 different food webs from a wide variety
  of ecosystems. We used three extinction scenarios: random, most
  connected, and least connected.
\item
  We found, as expected, that the predicted food webs were more robust
  than the observed food webs, which can be attributed to the higher
  connectance of the predicted food webs. On average, for every one unit
  of increase in connectance, we found that the food webs robustness
  increased by 0.52 units for the most connected species extinction
  scenario and by 0.04 units for the random species extinction
  scenarios. On the other hand, we saw no effect, on average, of
  increased connectance for the least connected species extinction
  scenario for all except two of the 12 food webs.
\item
  These results show that undersampling can lead to large underestimates
  of food web robustness that can be compensated for by filling in
  missing links with food web models. Nevertheless, increased
  connectance may contribute to lower dynamical stability, so it would
  be interesting to compare the dynamical stability of observed and
  predicted food webs and the topological stability that we have focused
  on.
\end{enumerate}
\end{abstract}

@@

Anthropogenic changes such as climate change and habitat destruction are
a threat to biodiversity and can lead to food web collapse (Ullah et
al., 2018). This food web collapse is due to the cascades of secondary
extinctions in a food web because of the primary loss of species due to
environmental change (Pimm et al., 2006; C. D. Thomas et al., 2004; J.
A. Thomas et al., 2004). An example of a secondary extinction is when a
consumer goes extinct because its sole resource species goes extinct.
Therefore, research focused on cascading secondary extinctions, also
known as `community viability analysis', has been performed extensively
in the past few decades to quantify how robust food webs are to species
extinction (Berg et al., 2011; Dunne et al., 2002a; Dunne \& Williams,
2009; Ebenman et al., 2004; Ebenman \& Jonsson, 2005). This research
revealed that the rate of collapse of a food web depends on its
structure and complexity (Dunne et al., 2002a; Dunne \& Williams, 2009).

Simulation of primary species loss has been conducted in observed food
webs and model food webs from terrestrial and aquatic ecosystems, where
robustness was measured in terms of secondary extinctions (Dunne et al.,
2002a; Dunne \& Williams, 2009). Primary species loss is considered to
be extinction of a species due to causes external to the food web,
e.g.~overharvesting (Koning \& McIntyre, 2021), introduction of invasive
species (David et al., 2017). A secondary extinction is an extinction
caused by a previous extinction (either primary or secondary). Numerous
studies have used topological criterion for assigning a secondary
extinction to a species, i.e., if extinction is of the last resource
species of a consumer species, then the consumer species suffers
secondary extinction (Dunne et al., 2002a; Dunne \& Williams, 2009;
Sol'e \& Montoya, 2001).

When there are few primary extinctions, and these cause many secondary
extinctions, a food web is said to be fragile or not robust. In
contrast, a food web is said to be robust when there are only a few
secondary extinctions. Several studies have shown that the robustness of
the food webs increases with food web connectance (Dunne et al., 2002b;
Dunne \& Williams, 2009). Additionally, these studies have demonstrated
that the removal (primary extinction) of the most connected species
causes considerably more secondary extinctions than the random removals
of species (Dunne et al., 2002b; Sol'e \& Montoya, 2001). Simulation
studies like these, which investigate the impact of primary extinctions
in a food web to quantify robustness based on its topological structure,
provide an alternate solution to canonical experiments in natural
ecosystems, which are not possible or very difficult to conduct (Dunne
\& Williams, 2009).

Along with quantifying food web robustness based on its topological
structure, studies such as Williams (2008), Brose et al. (2006), and
Martinez et al. (2006) have quantified robustness based on the abundance
dynamics of a food web. However, the topological approach of quantifying
a food web's robustness only requires the food web structure, whereas
the dynamical approach requires the food web structure and the temporal
dynamics of the abundance of species in that food web. For example,
Williams (2008) combined network structure models with bioenergetic
dynamics models to study the role of food web topology and nonlinear
dynamics on species coexistence in complex ecological networks.

A key assumption of the observed food webs is that they are very well
sampled, i.e.~all the links that in reality can occur are represented.
However, it is known that not all food webs are very well sampled and do
not represent all of the feeding links that can occur (Caron et al.,
2022; Jordano, 2016; Patonai \& Jord'an, 2017). Some rare trophic links
require more sampling effort than others, whereas some trophic links
remain unobserved because of biological constraints such as
spatio-temporal uncoupling, size or reward mismatching, foraging
constraints and physiological-biochemical constraints, which are
irrespective of sufficient sampling effort (Jordano, 2016). Previous
studies such as Caron et al. (2022) and Gupta et al. (2022) have shown
that the predicted food webs from these models usually have greater
connectance than the observed ones. One solution to compensate for
undersampling is using a food web model such as the Allometric Diet
Breadth Model (ADBM) (Petchey et al., 2008) to predict the missing links
and then measure the robustness of the predicted food web. The ADBM is a
mechanistic model constructed using foraging rules based on the body
sizes of prey and predator, where trophic interactions satisfying those
rules would be predicted by the model, which are perhaps not observed
because those interactions are rare. However, this solution is not
infallible, as it is likely that the food web model might still miss
some links and also may predict some links that could not, in fact
occur.

In our study, we investigate the topological robustness of the ADBM
predicted food webs and compare it to that of the observed food webs. We
expect the ADBM predicted food webs to be more robust than the observed
food webs, and for the greater robustness to be related to the amount by
which the ADBM predicts greater connectance compared to that of the
observed food webs. We do this by simulating primary species loss in 12
food webs predicted by the ADBM to quantify the secondary loss of
extinctions and compare the robustness of food webs predicted by the
ADBM to the robustness of the observed food webs. We use three different
approaches of primary species removal: (i) most connected species, (ii)
random species and (iii) least connected species to understand if the
outcome varies among these approaches.

@@

In the upcoming sections, we present a detailed account of the
implementation of the simulation of primary extinctions for three
different extinction scenarios on 12 food webs predicted by the ADBM
from a wide variety of ecosystems and compute the resultant secondary
extinctions. We then describe a robustness metric of those predicted
food webs and how we compare them to the properties of the food webs.

@@

The allometric diet breadth model (ADBM) is based on optimal foraging
theory, specifically the contingency model (MacArthur \& Pianka, 1966).
We chose this model because it is known to predict greater connectance
in the predicted food webs compared to the observed food webs (Gupta et
al., 2022). The ADBM predicts the set of prey species a consumer should
feed upon to maximise its rate of energy intake (Petchey et al., 2008).
The foraging variables used in the model are the energy content of prey,
handling times of the predator on prey, space clearance rate, i.e.~how
fast a predator searches space, and prey densities. Each of these
variables is derived from the allometric scaling relationship using the
body sizes of species. More details on the foraging rules defined in the
ADBM and ADBM's predictive power across different food webs can be found
in open access Petchey et al. (2008).

@@

The observed food webs that we fit the ADBM to belong to marine,
freshwater and terrestrial ecosystems (Table \ref{fig:tab_1}). We
considered these food webs because they belong to diverse ecosystems and
follow FAIR (Findable Accessible Interoperable Reusable) principles
(Wilkinson et al., 2016). These food webs contain primary producers,
herbivores, carnivores, parasites, and parasitoids and various feeding
interactions, including predation, herbivory, bacterivory, parasitism
and pathogenic. The observed connectance of these food webs varies from
0.03 to 0.24 (Links/Species\(^2\)), and the number of species varies
from 29 to 239. The goodness of fit of the ADBM's predictions depends on
the interaction types in the food webs. For example, those with
size-structured interactions, such as herbivory in aquatic ecosystems,
are better predicted compared to less size-structured ones, such as
parasitoids and terrestrial herbivory ones (Petchey et al., 2008).

@@

\caption{\label{tab:unnamed-chunk-1}\label{fig:tab_1}Information about the food webs predicted using the ADBM.}
\centering
\resizebox{\linewidth}{!}{
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}|>{\raggedright\arraybackslash}p{3cm}|l|>{\raggedright\arraybackslash}p{2cm}|>{\raggedright\arraybackslash}p{2cm}|>{\raggedright\arraybackslash}p{3cm}}
\hline
Common food web name (Original Publication) & Predation matrix source & General ecosystem & Number of species & Observed connectance (Links/Species$^2$) & 95\% prediction interval of predicted connectance  of the ADBM  (Gupta et al. 2022)\\
\hline
Benguela Pelagic (Yodzis 1998) & Brose et al. (2005) & Marine & 30 & 0.21 & 0.26 - 0.59\\
\hline
Broadstone Stream (taxonomic aggregation) (Woodward and Hildrew 2001; Woodward
et al. 2005) & Brose et al. (2005) & Freshwater & 29 & 0.19 & 0.18 - 0.72\\
\hline
Broom (Memmott et al. 2000) & Brose et al. (2005) & Terrestrial & 60 & 0.03 & 0.12 - 0.89\\
\hline
Capinteria (Lafferty et al. 2006) & Hechinger et al. (2011) & Marine (Salt Marsh) & 88 & 0.08 & 0.11 - 0.80\\
\hline
Caricaie Lakes (Cattin et al. 2004) & Brose et al. (2005) & Freshwater & 158 & 0.05 & 0.11 - 0.81\\
\hline
Grasslands (Dawah et al. 1995) & Brose et al. (2005) & Terrestrial & 65 & 0.03 & 0.03 - 0.44\\
\hline
Mill Stream (Ledger, Edwards, Woodward unpublished) & Brose et al. (2005) & Freshwater & 80 & 0.06 & 0.08 - 0.60\\
\hline
Skipwith Pond (Warren 1989) & Brose et al. (2005) & Freshwater & 71 & 0.07 & 0.17 - 0.90\\
\hline
Small Reef (Opitz 1996 Table 8.6.2) & Alyssa R. Cirtwill and Anna Eklöf (2018) & Marine (Reef) & 239 & 0.06 & 0.07 - 0.66\\
\hline
Tuesday Lake (Jonsson et al. 2005) & Brose et al. (2005) & Freshwater & 73 & 0.08 & 0.09 - 0.57\\
\hline
Ythan (Emmerson and Raffaelli 2004) & Alyssa R. Cirtwill and Anna Eklöf (2018) & Marine (Estuarine) & 85 & 0.04 & 0.13 - 0.84\\
\hline
Broadstone Stream (size aggregation) (Woodward
et al. 2010) & Guy Woodward (2021) & Freshwater & 29 & 0.24 & 0.25 - 0.47\\
\hline
\end{tabular}}
\end{table}
\end{landscape}
\restoregeometry

@@

The ADBM was parameterised using approximate Bayesian computation (ABC),
where a set of parameter values were sampled from a prior distribution.
Then, that set of parameter values was either accepted or rejected based
on how close the predicted food web was to the observed food web using
an accuracy metric -- true skill statistic (TSS). The accepted parameter
values formed a posterior distribution (Fig. 4 and S14-S25 in Gupta et
al. (2022)). Further, prediction intervals of the true skill statistic
and connectance of the predicted food webs were computed (Fig. 5 (a, b)
in Gupta et al. (2022)). In our study, we considered model predicted
food webs where the predicted connectance lay within the 95\% prediction
interval of all model predicted food webs. A detailed explanation of the
parameterisation method can be found in Gupta et al. (2022).

@@

We implemented the primary species removal method from Dunne \& Williams
(2009) by sequentially removing species using one of the three criteria:
removal of (i) the most connected species, (ii) the least connected
species and (iii) randomly chosen species. The most connected and least
connected criteria are based on species' degree (i.e.~the total number
of links to resources and from consumers). We considered these three
criteria because the random extinction scenario takes into account all
the theoretically possible extinction sequences of species that can
occur in a food web, while the extinction of the most connected species
and least connected species takes into account the two opposite extreme
scenarios. These extinction scenarios have been widely used in studying
species extinctions and the collapse of food webs and other networks
(Albert \& Barab'asi, 2002; J. Dunne et al., 2004; Dunne et al., 2002a;
Dunne \& Williams, 2009; Sol'e \& Montoya, 2001).

Given a primary removal of species in a food web, if any remaining
species lost all of their resource species, or any cannibalistic species
lost all of their resource species except the cannibalistic links, they
are removed from the web, and a secondary extinction was recorded.
Secondary extinctions may cause further secondary extinctions, which
were also checked for and recorded. Once no more secondary extinctions
occurred, then another primary extinction was made of the next
appropriate species depending on the extinction scenario. This process
was carried out until all the species were extinct from the food web.

The robustness (R) of a food web was defined as the proportion of
species subjected to primary removals resulting in extinction (primary
and secondary extinctions) of some specified proportion of the species.
In our study, we use \(R_{50}\), the number of primary extinctions
divided by the total number of species, which results in at least 50\%
of total species loss (J. Dunne et al., 2004; Dunne et al., 2002a; Dunne
\& Williams, 2009; Jonsson et al., 2015). Therefore, if primary
extinctions never cause any secondary extinctions, the food web is
maximally robust and (\(R_{50} = 0.50\)). Whereas in a minimally robust
community (\(R_{50} = 1/S\), where \(S\) is the number of species), the
first primary extinction causes a cascade of secondary extinctions of at
least nearly half of the species in the food web (i.e.~at least
\(S/2 - 1\)).

@@

First, we simulated primary species loss in food webs predicted by the
ADBM which had the maximum true skill statistics and compared it to
primary species loss in observed food webs. Second, to take into account
the uncertainty in robustness in the ADBM predicted food webs, we
simulated primary species loss and thereby computed robustness for all
the ADBM predicted food webs corresponding to the 95\% prediction
interval of the predicted connectance. Furthermore, in the case of the
random extinction scenario, we simulated 1000 random extinction
sequences in a single ADBM predicted food web.

\hypertarget{analysis}{%
\subsection{Analysis}\label{analysis}}

In the random extinction scenario, we computed robustness \(R_{50}\) for
all 1000 independent random extinction sequences and calculated the
median as a summary statistics to quantify the average robustness of a
single food web to random extinction. To quantify the effect of
undersampling, i.e.~greater connectance of connectance, we compute the
ratio of the difference in normalised robustness between the ADBM
predicted food webs and observed food webs to the difference in their
normalised connectance, where normalisation was performed by dividing
the variables by their maximum possible values (i.e.~0.5 for \(R_{50}\)
and 1 for connectance). However, we did not perform any statistical
significance test because we work with simulated food webs and
therefore, the p-values of these tests are influenced by the number of
model simulations (White et al., 2014).

@@

We first present the secondary extinction curves of the ADBM predicted
food webs which had the maximum true skill statistics and of the
observed food webs for 12 food webs under three different extinction
scenarios. We then compare the robustness of all the ADBM predicted food
webs within the 95\% prediction interval against that of the observed
food webs to take into account uncertainty in the robustness across food
webs predictions. Finally, we quantify the effect of the difference in
their connectance on the difference in their robustness estimates.

@@

In the most connected extinction scenario, the cumulative secondary
extinction curves started to rise steeply in the observed food webs at a
lower number of primary species removal as compared to the ADBM
predicted food webs for nine food webs (Fig. \ref{fig:fig_r1} (a, c, d,
e, f, g, h, j, l)). However, there was a higher number of cumulative
secondary extinctions occurring in the ADBM predicted food webs when
compared to that of the observed food webs at a high number of primary
species removal in five food webs (Fig. \ref{fig:fig_r1} (a, f, g, h,
k)). In the Skipwith Pond food web, there were no secondary extinctions
for any number of primary removal of species (Fig. \ref{fig:fig_r1}
(i)), whereas in the Broadstone Stream (taxonomic aggregation) food web
the same was true only for the observed food web but in the ADBM
predicted food web there was a steep rise in the cumulative secondary
extinctions (Fig. \ref{fig:fig_r1} (b)).

@@

\caption{\label{fig:fig_r1} Cumulative secondary extinctions of species resulting from the primary removals of the \textbf{most connected species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species.}\label{fig:unnamed-chunk-2}
\end{figure}

In Fig. \ref{fig:fig_r2}, we present the cumulative secondary
extinctions in the ADBM predicted and the observed food webs for five
(out of 1000) independent random extinction sequences to show example
variation caused by different random primary extinction orders. The
secondary extinction curves of the ADBM predicted food webs were steeper
than that of the observed food webs, i.e.~primary removal of some
species in an extinction sequence can lead to the complete collapse of
the remaining food web in the ADBM predicted food webs.

@@

\caption{\label{fig:fig_r2} Cumulative secondary extinctions of species resulting from the primary removals of \textbf{random species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs for five (out ot 1000) independent random extinction sequences. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species.}\label{fig:unnamed-chunk-3}
\end{figure}

Compared to the most connected and random extinction scenarios, there
were fewer secondary extinctions in the least connected extinction
scenario, and therefore the secondary extinction curves were flat for
many of the food webs (Fig. \ref{fig:fig_r3}). In some of the food webs,
the extinction curves of the ADBM predicted food webs overlapped with
the observed food webs (Fig. \ref{fig:fig_r3} (b, c, g, h, i, k, l)\}).
In five of the food webs, a very high number of secondary extinctions
occurred at a very low number of primary species removal (Fig.
\ref{fig:fig_r3} (a, d, e, f, j)).
